### 2020-12-06 paper/localpsf.tex
edits to poiwwon weighting functions section

\section{Interpolating impulse response samples}

\section{Poisson PSF interpolation}
\label{sec:weighting_functions}

We form an interpolant, $\widetilde{\varphi}_x$, of the point spread function (PSF), $\varphi_x$, as follows:
\begin{equation*}
	\varphi_x(y) \approx \widetilde{\varphi}_x(y) := \sum_{i=1}^q w_i(x)\varphi_{x_i}(y).
\end{equation*}
The functions $y \mapsto \varphi_{x_i}(y)$, are samples of the PSF at the sample points $x_i$, $i=1,\dots,q$. The functions $w_i(x)$, are smooth spatially varying weighting functions that solve the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$. We refer to $\widetilde{\varphi}_x$ as the \emph{Poisson interpolant} of $\varphi_x$.

Let $u_y$ denote the function $u_y(x) := \widetilde{\varphi}_x(y)$ for an arbitrary but fixed value of $y$. It is straightforward to show that $u_y$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{u_y} &\quad \frac{1}{2}\norm{\Delta u_y}_{L^2(\Omega)}^2 \\
\text{such that} &\quad u_y(x_i) = \varphi_{x_i}(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
This follows from linearity of the KKT system associated with \eqref{eq:wi_optimization_problem}, which in turn follows from the fact that the the objective function in \eqref{eq:wi_optimization_problem} is quadratic, and the constraint is linear.
Since the Laplacian measures local deviation from flatness [cite Jacobsen], from \eqref{eq:uy_optimization_problem} we see that the interpolant $x \mapsto u_y(x) = \widetilde{\varphi}_x(y)$ is the `flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and interpolates the data $\varphi_{x_i}(y)$.

@@

In Section SEC, we will show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$. Briefly, constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed efficiently with multigrid. Radial basis function interpolation of $\varphi_x(y)$ is cheaper per sample point because radial basis functions are given by analytic formulas. However, the cost of these Poisson PDE solves is typically cheap compared to the cost of computing the impulse response samples, $\varphi_{x_i}$. Computing the impulse response samples, $\varphi_{x_i}$, requires applying the Hessian to vectors, which in turn requires performing PDE solves of the linearized forward and ajoint equations for the inverse problem. These forward and adjoint PDE solves are typically much more expensive than Poisson PDE solves, because the Poisson equation is well-studied and easy to solve with multigrid.

@@

>>>>

\section{Poisson interpolation weighting functions}
\label{sec:weighting_functions}

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$. Since the Laplacian measures local deviation from flatness [cite Jacobsen], the weighting function $w_i$ is the ``flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and takes the value $1$ at point $x_i$, and zero at the points $x_j$, $j \neq i$. In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense.

\begin{thm}
	\label{thm:smoothest_interpolant}
Let $u$ be the function
\begin{equation*}
u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
\end{equation*}
for an arbitrary but fixed value of $y$. Then $u$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{u} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}
\begin{proof}
This follows from linearity of the KKT system associated with \eqref{eq:wi_optimization_problem}, which in turn follows from the fact that the the objective function in \eqref{eq:wi_optimization_problem} is quadratic, and the constraint is linear.
\end{proof}

%Since the Laplacian measures local deviation from flatness [cite Jacobsen], from \eqref{eq:uy_optimization_problem} we see that the interpolant $x \mapsto u_y(x) = \widetilde{\varphi}_x(y)$ is the ``flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and interpolates the data $\varphi_{x_i}(y)$.

@@

@@

In this section, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$. Briefly, constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All the existing weighting functions, $\{w_i\}_{i=1}^q$, will change if a new point, $x_{q+1}$, is added. We present an efficient incremental algorithm for updating these weighting functions when a new point is added, in which we only need to perform two new Poisson PDE solves (rather than $2(q+1)$ new PDE solves).

%Radial basis function interpolation of $\varphi_x(y)$ is cheaper per sample point because radial basis functions are given by analytic formulas. However, the cost of these Poisson PDE solves is typically cheap compared to the cost of computing the impulse response samples, $\varphi_{x_i}$. Computing the impulse response samples, $\varphi_{x_i}$, requires applying the Hessian to vectors, which in turn requires performing PDE solves of the linearized forward and ajoint equations for the inverse problem. These forward and adjoint PDE solves are typically much more expensive than Poisson PDE solves, because the Poisson equation is well-studied and easy to solve with multigrid.
